Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces

### 2025-03-20 paper/main.tex
Universal commits presents...

\paragraph{Data Preparation.}
We use the MNIST dataset, which consists of $60{,}000$ training images and $10{,}000$ test
images of handwritten digits (0--9). Each image is $28\times28$ pixels, grayscale.

@@

\begin{enumerate}
  \item \textbf{Sampling Directions.} Generate $K$ independent directions $\mathbf{d}_1, \dots, \mathbf{d}_M$
        from $\mathcal{N}(0,I_d)$.
  \item \textbf{Loss Evaluation.} For each direction $\mathbf{d}_t$, evaluate the loss at
        $\mathbf{w}^* + \mathbf{d}_t$ and
        $$\left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 $$
  \item \textbf{Averaging.} Average these values values over $k = 1, \dots, K$ to approximate $\Delta_k$.
\end{enumerate}

@@

\paragraph{Rationale}
From prior work \cite{sagun2018empirical}, we know that only a few Hessian eigenvectors (those associated with the largest
eigenvalues) tend to capture the most significant curvature directions. By projecting onto these principal directions,
we can more precisely measure how the landscape the loss surface changes as $k$ increases.

\paragraph{Procedure}
\begin{enumerate}
  \item \textbf{Hessian approximation.} For each trained model, approximate
        the Hessian $\mathbf{H}$ or its dominant eigenvalues/eigenvectors. Techniques ???
  \item \textbf{Eigen-decomposition.} Identify the top $d$ eigenvectors $\{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ with the
        largest eigenvalues.
  \item \textbf{Subspace exploration.} For each model, calculate a $\Delta$ fuction via the same way as in the second
        experiment (using Monte-Carlo) and appy visualization/.
\end{enumerate}

\paragraph{Goal and Expected Outcome}
\begin{itemize}
  \item \textbf{Refined understanding.} By focusing on the directions of largest curvature, we expect a clearer
        measure of how the landscape of the loss surface changes with additional data.
  \item \textbf{Identifying the threshold.} This approach should yield a more accurate estimate of the point at which the
        landscape effectively stabilizes, thus helping to pinpoint a minimum viable dataset size.
  \item \textbf{Connect with theoretical results.} By applying experiments we check our calculations above on the reality.
\end{itemize}

\begin{figure}[!htbp]
  \hspace*{-0.7cm}
  \subfloat{\includegraphics[width=1.15\textwidth]{img/D_32.pdf}}\\
  \caption{$\Delta$-function visualization. `Sigma` parameter refers to the variance of sampled gaussian vectors, `dim' 
  refers to the rank of the subspace.}
  \label{fig:base_delta}
\end{figure}

>>>>

\paragraph{Data.}
We use the MNIST dataset, which consists of $60{,}000$ training images and $10{,}000$ test
images of handwritten digits (0--9). Each image is $28\times28$ pixels, grayscale.

@@

\begin{enumerate}
  \item \textbf{Sampling Directions.} Generate $K$ independent directions $\mathbf{d}_1, \dots, \mathbf{d}_M$
        from $\mathcal{N}(0,I_d)$.
  \item \textbf{Decreasing multidimension.} Project the network’s parameters onto the subspace they span.
  \item \textbf{Loss Evaluation.} For each direction $\mathbf{d}_t$, evaluate the loss at
        $\mathbf{w}^* + \mathbf{d}_t$ and
        $$\left( \mathcal{L}_{k}(\bw + \mathbf{d}_t) - \mathcal{L}_{k-1}(\bw + \mathbf{d}_t) \right)^2 $$
  \item \textbf{Averaging.} Average these values values over $k = 1, \dots, K$ to approximate $\Delta_k$.
\end{enumerate}

@@

From prior work \cite{sagun2018empirical}, we know that only a few Hessian eigenvectors (those associated with the largest
eigenvalues) tend to capture the most significant curvature directions. By projecting onto these principal directions,
we can more precisely measure how the landscape the loss surface changes as $k$ increases.

\paragraph{Procedure}
\begin{enumerate}
  \item \textbf{Hessian approximation.} For each trained model, approximate
        the Hessian $\mathbf{H}$ or its dominant eigenvalues/eigenvectors. Techniques ???
  \item \textbf{Eigen-decomposition.} Identify the top $d$ eigenvectors $\{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ with the
        largest eigenvalues.
  \item \textbf{Calculation and analization.} For each model, calculate a $\Delta$ fuction via the same way as in the second
        experiment (using Monte-Carlo) and visualize via plots.
\end{enumerate}

\paragraph{Goal and Expected Outcome}
\begin{itemize}
  \item By focusing on the directions of largest curvature, we expect a clearer
        measure of how the landscape of the loss surface changes with additional data.
  \item This approach should yield a more accurate estimate of the point at which the
        landscape effectively stabilizes, thus helping to pinpoint a minimum viable dataset size.
  \item By applying experiments we check our calculations above on the reality.
\end{itemize}

\begin{figure}[!htbp]
  \centering
  \subfloat{\includegraphics[width=1\textwidth]{img/D_32.pdf}}\\
  \caption{$\Delta$-function visualization. `Sigma` parameter refers to the variance of sampled gaussian vectors, `dim'
    refers to the rank of the subspace.}
  \label{fig:base_delta}
\end{figure}
